The Design and Implementation of the Synchronous Language \CEU

### 2016-04-01 paper.tex
...

Esterel defines time as a discrete sequence of logical unit instants or 
``ticks''.
At each tick, the program reacts to an arbitrary number of simultaneous input 
events, depending on external stimuli the environment provides.
In contrast, \CEU defines time as a discrete sequence of reactions to
unique input events.
At each input event, which constitutes a logical unit of time, the program 
reacts exclusively to it.
The execution of a program in \CEU is as follows~\cite{ceu.sensys13}:
%
\begin{enumerate}
\item The program initiates the ``boot reaction'' in a single trail (but 
      parallel constructs may create new trails).
\item Active trails execute until they await or terminate.
      This step is named a \emph{reaction chain}, and always runs in bounded 
      time.
\item The program goes idle and the environment takes control.
\item On the occurrence of a new external input event, the environment awakes 
      \emph{all} trails awaiting that event.
      It then goes to step 2.
\end{enumerate}
%
A program must react to an occurring event completely before handling the next.
%
Based on the synchronous hypothesis, a program takes no time on step 2 and is 
always idle on step 3.
In practice, if a new external input event occurs while a reaction chain is 
running (step 2), it is enqueued to occur in the next reaction.

@@

@@

@@

@@

To deal with the second limitation, which can be critical in the presence of 
lengthy computations, we can consider the following approaches:
(1) manually placing \code{pause} statements in unbounded loops;
(2) integrating \CEU with a preemptive system.
%
The first option requires the lengthy operations to be rewritten in \CEU using
\code{pause} statements so that other trails can be interleaved with them.
This option is the one recommended in many related work that provide a similar 
cooperative primitive (e.g. \code{pause}~\cite{esterel.primer}, 
\code{PT\_YIELD}~\cite{wsn.protothreads}, \code{yield}~\cite{wsn.sol}, 
\code{post}~\cite{wsn.nesc}).
%
Considering the second option, \CEU and preemptive threads are not mutually 
exclusive.
% and the second option is appealing to be explored in a future work.
For instance, TOSThreads~\cite{wsn.tosthreads} proposes a message-based 
integration with \emph{nesC} that is safe and matches the semantics of \CEU 
external events.
\end{comment}

@@

% CONTRIBS
    - queue in the semantics
    - permeates all aspects
    - As far as we know, \CEU is the first Esterel-based language to HAVE?
      queue-based/unique in the semantics of the language and not in the
      library
    - stack
        - deterministic intra-reaction execution
    - imposed intra-reaction execution
        - for inter-reaction commm
        - for intra-reaction commm
        - for stmts

We presented the design and implementation of \CEU, a synchronous reactive 
language inspired by Esterel targeting constrained embedded systems.

@@

\begin{comment}
We are also extending and applying \CEU in more dynamic domains, such as GUIs 
and games~\cite{ceu.mod15}, by relaxing some restrictions in the language.
In particular, programs in these domains typically require unbounded memory and 
execution time, such as for dealing with dynamic collections.
We provide unsafe annotations to circumvent the rigid semantics of \CEU, 
requiring them to be explicit and trackable at the same time.
\end{comment}

>>>>

Esterel defines time as a discrete sequence of logical unit instants or 
``ticks''.
At each tick, the program reacts to an arbitrary number of simultaneous input 
events, depending on external stimuli the environment provides.
In contrast, \CEU defines time as a discrete sequence of reactions to
unique input events.
At each input event, which constitutes a logical unit of time, the program 
reacts exclusively to it.
The event-triggered execution of a program in \CEU is as 
follows~\cite{ceu.sensys13}:
%
\begin{enumerate}
\item The program initiates the ``boot reaction'' in a single trail (but 
      parallel constructs may create new trails).
\item Active trails execute until they await or terminate.
      This step is named a \emph{reaction chain}, and always runs in bounded 
      time.
\item The program goes idle and the environment takes control.
\item On the occurrence of a new external input event, the environment awakes 
      \emph{all} trails awaiting that event.
      It then goes to step 2.
\end{enumerate}
%
A program must react to an occurring event completely before handling the next.
%
Based on the synchronous hypothesis, a program takes no time on step 2 and is 
always idle on step 3.
In practice, if a new external input event occurs while a reaction chain is 
running (step 2), it is enqueued to occur in the next reaction.

@@

A limitation of event-triggered execution is that all program behavior must be 
purely reactive, given that no code can execute in the absence of inputs.
%
Tick-triggered execution allows for active behavior, since code can execute 
regulary on every tick.
%
Although it supports asynchronous active execution~\cite{ceu.tr}, we still 
consider \CEU as purely reactive since its synchronous core cannot express 
active behavior.

@@

\subsection{Summary}

Contributions:
%
\begin{itemize}
\item \textbf{Event-triggered notion of time bound to the semantics of the 
language}, which allows ....
\item \textbf{Deterministic and fine-grained intra-reaction execution and 
communication}, lexical, stack-based internal events
\item \textbf{Static temporal analysis for concurrent statements}, which
      prevent logical race conditions on shared variables and system calls.
\item \textbf{Safe integration with \C}, with enforced resource clean up.
\item \textbf{First-class timers} with dedicated syntax and automatic 
synchronization.
\end{itemize}

Limitations:
%
\begin{itemize}
\item purely reactive execution,
\item bounded-only reactions,
\item sequential
\end{itemize}
%
We believe that a simple-tractable synchronous core
and extend it with memory-isolated parallelizabe asynchronous execution.

    - permeates all aspects
    - As far as we know, \CEU is the first Esterel-based language to HAVE?
      queue-based/unique in the semantics of the language and not in the
      library
    - stack
        - deterministic intra-reaction execution
    - imposed intra-reaction execution
        - for inter-reaction commm
        - for intra-reaction commm
        - for stmts

@@

TODO: comparison with Esterel

@@

To deal with the second limitation, which can be critical in the presence of 
lengthy computations, we can consider the following approaches:
(1) manually placing \code{pause} statements in unbounded loops;
(2) integrating \CEU with a preemptive system.
%
The first option requires the lengthy operations to be rewritten in \CEU using
\code{pause} statements so that other trails can be interleaved with them.
This option is the one recommended in many related work that provide a similar 
cooperative primitive (e.g. \code{pause}~\cite{esterel.primer}, 
\code{PT\_YIELD}~\cite{wsn.protothreads}, \code{yield}~\cite{wsn.sol}, 
\code{post}~\cite{wsn.nesc}).
%
Considering the second option, \CEU and preemptive threads are not mutually 
exclusive.
% and the second option is appealing to be explored in a future work.
For instance, TOSThreads~\cite{wsn.tosthreads} proposes a message-based 
integration with \emph{nesC} that is safe and matches the semantics of \CEU 
external events.

TODO: search "limitation" in the paper and condense all

A limitation of event-triggered execution is that all program behavior must be 
purely reactive, given that no code can execute in the absence of inputs.
%
Tick-triggered execution allows for active behavior, since code can execute 
regulary on every tick.
%
Although it supports asynchronous active execution~\cite{ceu.tr}, we still 
consider \CEU as purely reactive since its synchronous core cannot express 
active behavior.

\end{comment}

@@

We presented the design, semantics, and implementation of \CEU, a synchronous 
reactive language inspired by Esterel targeting constrained embedded systems.

@@
